### 2015-03-09 article/abstract.tex
article initial commit

Nuclear receptors (NRs) are important signaling mediators in cells due to their ability to act as receptors and transcription factors. These molecules are activated by ligands and regulate critical functions in cell control, inflammation, fibrosis and tumor formation. Nuclear receptors act on the metabolism and signaling of cells by changing the expression of target genes. Here, we will investigate the knockout phenotypes and genetic variation of mouse NRs. We will first create an assembly of all known mouse SNPs in the vicinity of mouse NR genes (+-1Mbp up-or downstream). In a second step, phenotype information for genetic knockouts and genetic variation data will be included. Knockout phenotypes are available from the MGI database, while the Mouse Phenome Database provides SNPs from various mouse strains, which can be correlated to extreme phenotypes, measured in these mouse strains. The goal of this analysis is to find NR SNPs in mice that influence changes in biological parameters such as body weight, body fat and other phenotypic traits. Furthermore, we will couple these findings to phenotypes observed in mice with a targeted or spontaneous mutation of the nuclear receptor and thus provide additional indication for a putative functionality of the investigated SNPs.